DATA 621 - Business Analytics and Data Mining

### 2020-09-24 HW2/HW2-Group2.rmd
Part 9

# 2. Confusion Matrix
A [confusion matrix](http://www.saedsayad.com/model_evaluation_c.htm) is a
format for summarizing the performance of a classification algorithm. It shows
the number of correct and incorrect predictions made by the classification model
and compares it to the actual outcomes in the data. The matrix is
\(N \times N\), where \(N\) is the number of target values. The following table
displays a \(2\times 2\) confusion matrix for two classes---positive and
negative. The actual confusion matrix are the 4 cells containing `a`, `b`, `c`,
and `d`. The enhancments around the matrix display the definitions for many
common classification metrics.

@@

>>>>

# 2. Confusion Matrix
A [confusion matrix](http://www.saedsayad.com/model_evaluation_c.htm) is a
format for summarizing the performance of a classification algorithm. It shows
the number of correct and incorrect predictions made by the classification model
and compares it to the actual outcomes in the data. The matrix is
\(N \times N\), where \(N\) is the number of target values. The following table
displays a \(2\times 2\) confusion matrix for two classes---positive and
negative. The actual confusion matrix are the 4 cells containing `a`, `b`, `c`,
and `d`. The enhancements around the matrix display the definitions for many
common classification metrics.

@@

# 9. Bounds on \(F_1\)
## Proof 1
We can prove this by decomposing the metrics into their component parts as
defined above.
\[
\begin{align}
F_{1:numerator} &= 2 \cdot P \cdot S\\
&= 2 \cdot \frac{TP}{TP + FP} \cdot \frac{TP}{TP + FN}\\
&= \frac{2\cdot(TP)^2}{(TP+FP)(TP+FN)}\\
F_{1:denominator} &= P + S\\
&= \frac{TP}{TP + FP} + \frac{TP}{TP + FN}\\
&= \frac{TP(TP + FN) + TP(TP+FP)}{(TP+FP)(TP+FN)}\\
&= \frac{(TP)^2 + TP\cdot FN + (TP)^2 + TP\cdot FP}{(TP+FP)(TP+FN)}\\
&= \frac{2\cdot(TP)^2 + TP\cdot (FN + FP)}{(TP+FP)(TP+FN)}\\\
F_1 &= \frac{F_{1:numerator}}{F_{1:denominator}}\\
&= \frac{2\cdot(TP)^2}{2\cdot(TP)^2 + TP\cdot (FN + FP)}\\
&= \frac{2\cdot TP}{2\cdot TP + FN + FP}
\end{align}
\]

Now this last equation ***must*** be bounded between 0 and 1. Consider that TP
is a non-negative integer. Also, the sum of TP, FP, TN, and FN must equal the
number of observations. Excluding the degenerate case where the dataset contains
only negative classes and they were all scored properly, the denominator will
always be positive. When the dataset consists of only positive classes and they
are all coded properly, we have \(F_1 = 1\). This is also why there isn't too
much to worry about with the all-negative case, as one can simply switch the
labels and side-step the division by zero error. In every other case where TP
is strictly less than the number of observations, there must be entries in
FN or FP. This causes the fraction to be less than one. In the case where no
positive entries are coded properly, then the numerator is 0 and the denominator
is > 0, so the \(F_1\) score is 0. As the building blocks all represent counts,
they cannot be negative, so the score can never fall below 0. ***QED***

## Proof 2
Another proof relies on the assumption that the sensitivity and precision
themselves are constrained to \([0, 1]\). Given those assumptions, it ***must***
be true that \(2ab < a + b\). This is a corollary of the
[*AM-GM inequality*](https://en.wikipedia.org/wiki/Inequality_of_arithmetic_and_geometric_means)
since if \(a, b \in [0, 1]\) then \(\sqrt{ab} > ab\) and this inequality is
strict. Thus the numerator must be strictly less than the denominator. Both
numerator and denominator are bounded by \([0, 1]\) as per the initial
assumptions. Therefore, excepting the degenerate case where both are 0,
this completed the proof the fraction too is bounded by \([0, 1]\). ***QED***
